---
title: 'CMTH 642 Data Analytics: Advanced Methods'
output:
  word_document: default
  html_document: default
  pdf_document: default
---
<center> <h1> Assignment 3 (10%) </h1> </center>
<center>  <h3> [Antonio Andrade] </h2> </center>
<center> <h3> [CMTH-642 & 501017196] </h2> </center>
---


#### 1. Import to R the following file: http://archive.ics.uci.edu/ml/machine-learning-databases/wine-quality/winequality-white.csv (The dataset is related to white Portuguese "Vinho Verde" wine. For more info: https://archive.ics.uci.edu/ml/datasets/Wine+Quality) (3 points)
```{r}
wine <- read.csv("~/CMHT642 - DHA/Assignment 3/winequality-white.csv", sep=";")
```

#### 2. Check the datatypes of the attributes. (3 points)
```{r}
str(wine)
```
  
#### 3. Are there any missing values in the dataset? (4 points)
```{r}
sum(is.na(wine))
table(is.na(wine))
#There are no missing values in the dataset
```

#### 4. What is the correlation between the attributes other than Quality? (10 points)
```{r}
cor(wine)
library(PerformanceAnalytics)
winesub <- wine[,c(1:11)]
chart.Correlation(winesub, histogram = TRUE, method = "pearson")
# All the other attributes are numeric class. The Quality attribute is defined as an integer.
```

#### 5. Graph the frequency distribution of wine quality by using Quality. (10 points)
```{r}
hist(wine$quality, main = "Wine quality distribution", xlab = "Quality")
```

#### 6.	Reduce the levels of rating for quality to three levels as high, medium and low. Assign the levels of 3 and 4 to level 0; 5 and 6 to level 1; and 7,8 and 9 to level 2. (10 points)
```{r}
library(forcats)
wine[,12] <- factor(wine[,12])
class(wine$quality)
quality_reduced <- fct_collapse(wine$quality, Low = c("3", "4"), Medium = c("5", "6"), High = c("7", "8", "9"))
fct_count(quality_reduced)
wine[,"quality_reduced"] <- quality_reduced
wine <- wine[-12]
```

#### 7. Normalize the data set by using the following function: (12 points)
```{r}
normalize <- function(x){
  return ((x - min(x)) / (max(x) - min(x)))
}
```

```{r}
wine_n <- as.data.frame(lapply(wine[1:11], normalize))
wine_n <- cbind(quality_reduced, wine_n)
```

#### 8. Divide the dataset to training and test sets. (12 points)
```{r}
set.seed(550)
index <- sample(1:nrow(wine_n), 0.70*nrow(wine_n))
wine_train <- wine_n[index,]
wine_test <- wine_n[-index,]
```

#### 9. Use the KNN algorithm to predict the quality of wine using its attributes. (12 points)
```{r}
library(class)
library(gmodels)
wine_train_labels <- wine_train[,1] 
wine_test_labels <- wine_test[,1]

wine_test_pred <- knn(train = wine_train[,2:11], test = wine_test[,2:11], cl = wine_train[,1], k = 10)

```

#### 10. Display the confusion matrix to evaluate the model performance. (12 points)
```{r}
cm <- table(Actual=wine_test_labels, Predicted=wine_test_pred)
cm
```

#### 11. Evaluate the model performance by computing Accuracy, Sensitivity and Specificity. (12 points)
```{r}
accuracy <- (cm[1,1] + cm[2,2] + cm[3,3])/(cm[1,1] + cm[1,2] + cm[1,3] + cm[2,1] + cm[2,2] + cm[2,3] + cm[3,1] + cm[3,2] + cm[3,3])
accuracy

sensitivityLow <- cm[1,1]/(cm[1,1] + cm[1,2] + cm[1,3])
sensitivityLow

sensitivityMedium <- cm[2,2]/(cm[2,1] + cm[2,2] + cm[2,3])
sensitivityMedium

sensitivityHigh <- cm[3,3]/(cm[3,1] + cm[3,2] + cm[3,3])
sensitivityHigh

specificityLow <- (cm[2,2] + cm[2,3] + cm[3,2] + cm[3,3])/(cm[2,1] + cm[2,2] + cm[2,3] + cm[3,1] + cm[3,2] + cm[3,3]) 
specificityLow

specificityMedium <- (cm[1,1] + cm[1,3] + cm[3,1] + cm[3,3])/(cm[1,1] + cm[1,2] + cm[1,3] + cm[3,1] + cm[3,2] + cm[3,3])
specificityMedium

specificityHigh <- (cm[1,1] + cm[1,2] + cm[2,1] + cm[2,2])/(cm[1,1] + cm[1,2] + cm[1,3] + cm[2,1] + cm[2,2] + cm[2,3])
specificityHigh
```

This is the end of Assignment 3

Ceni Babaoglu, PhD